\section{November 7}
\subsection{Proof Systems}
By Russel's Paradox, sets "cannot" be non-wellfounded.
\begin{corollary}
  There is no "proof system" in Number Theory
  \begin{itemize}
    \item Soundness: No false statement $\phi$ has a valid proof $\pi$.
    \item Completeness: Every trye statement $\phi$ has a valid proof $\pi$.
    \item Decidability: Can decide if $\pi$ is a valid proof for $\phi$
  \end{itemize}
\end{corollary}
We generally give up soundness for completeness.
\begin{proof}
  If there was a proof system for number theory, then we could decide number theory.
  \begin{itemize}
    \item Given $\phi$ try all valid strings $\pi$ one-by-one:
    \begin{itemize}
      \item Check if $\pi$ is a valid proof of $\phi$. If so, accept.
      \item Check if $\pi$ is a valid proof of $\neg \phi$. If so, reject.
    \end{itemize}
  \end{itemize}
\end{proof}

\subsection{Gödel's Sentence}
We want to construct a statement in a proof system that is true and unprovable.
\\
Take proposition p(x) = "p(x) is unprovable."
\begin{enumerate}
  \item Proving p(x) is true: $\neg p(x) \implies p(x) \text{ is provable } \implies p(x)$ by the property of soundedness in proof systems. This is clearly a contradiction.
  \item If p(x) is true, it must also be unprovable.
\end{enumerate}
Showing that a proposition can describe itself.
\\
$\Phi^*(n):$
\begin{enumerate}
  \item parse n as $\langle \phi \rangle$
  \item True, if $\nexists$ proof for $\phi(n)$
\end{enumerate}
$\phi^*(n) := \nexists$ proof for $\phi(n)$ where $\langle \phi \rangle = n$.
\\
Consider $\phi^*(\langle \phi^* \rangle)$
\begin{enumerate}
  \item Proving $\phi^*(\langle \phi^* \rangle)$ is true: Suppose not, $\implies \exists$ proof for $\phi^*(\langle \phi^* \rangle)$. By soundedness, we then know that $\phi^*(\langle \phi^* \rangle)$ is true.
  \item Proving $\phi^*(\langle \phi^* \rangle)$ is unprovable: Suppose not, $\implies \phi^*(\langle \phi^* \rangle)$ is true which results in a contradiction.
\end{enumerate}
We can use diagonalization to build a matrix where each row represents the mathematical equation $phi$ and each column represents a natural number n. Each cell represents if there exists a proof for $\phi(n)$.