\section{References and Aliasing}
In OCaml, reference cells are type-invariant after they are created. 
\\
We can think as references as a pointer to a memory cell on the heap. Aliasing just closes the value with the same address.
\\
When comparing two references with (=), the operator checks if their values are the same. We might want to treat them differently.
\begin{lstlisting}
let f (x : int ref) (y : int ref) =
  y := 0;
  !x
\end{lstlisting}
This function is intended to:
\begin{itemize}
  \item return contents of x
  \item modify y
  \item don't modify x
\end{itemize}
But when we call \textit{let $\_$ = f x x} this will reset x which we don't want. This is a natural solution.
\begin{lstlisting}
let f (x : int ref) (y : int ref) int option =
  if x = y
  then None
  else (y := 0; Some(!x))
\end{lstlisting}
This code is sound as it follows the requirements but will return None if we pass two different references with the same value. 
\\
Instead of structural equality, we need physical equality (==) to check if the addresses are the same.
\begin{lstlisting}
let f (x : int ref) (y : int ref) int option =
  if x == y
  then None
  else (y := 0; Some(!x))
\end{lstlisting}
It becomes challenging to handle these things using logic.
\begin{itemize}
  \item References may be mutated over test cases
  \item You must reason about the addresses and values
  \item Side effect that happens over time.
\end{itemize}
Monads??
\\
In order to test the above function, we need to make a wrapper function with an inputed edge case.
\begin{lstlisting}
let fwrapper (x : int) (y : int) (alias : bool) 
: (int option * int * int) =
  let (xref, yref) =
    let xref = ref x in 
    if alias
    then (xref, xref)
    else 
      let yref = ref y in
      (xref, yref)
  in
  let o = (f xref yref) in
  (o, !xref, !yref)
\end{lstlisting}
The property now becomes
\begin{lstlisting}
let prop_mut (ix, iy, alias) : bool =
  let (o, fx, fy) = fwrapper ix iy alias in
  if alias
  then o = None && ix = fx && iy = fy
  else 
    match o with
    | None -> false
    | Some v -> v = fx && fy = 0 && fx = ix
\end{lstlisting}
We still have an issue with the side effect over time problem which we might want to test. It could be important to ensure that we are maintaining invariants dynamically.